\thetitle

### 2021-05-07 writing/introduction.tex
Touch up previous edits

The paper's {\it second contribution\/} is a set of {\em results from
applying the evaluation method to three distinct blame assignment
strategies in approximately 72,000 different scenarios\/} (see
sec.~\ref{sec:results}): (1)~{\it Transient\/}, i.e. Reticulated's
tracking of typed/untyped boundaries~\citep{vss-popl-2017}; (2)~{\it
Natural\/}, i.e. Typed Racket's use of a higher-order contract system and
its blame assignment~\citep{ff-icfp-2002}; and (3) {\it Erasure\/}, i.e.
the approach of industrial systems, which forgo blame in favor of error
messages from the safety checks of the underlying language. The results
(see sec.~\ref{sec:discussion}) are at least somewhat surprising.  In
principle they validate the conjectures behind the work of theoreticians.
First, a blame assignment strategy helps with the search for
impedance mismatches between the specified types and the behavior of
untyped components.  Second, Natural's wrapper-based blame tracking is
more useful than Transient's ``collective blame'' tracking algorithm,
which in turn is superior to Erasure. {\em But\/}, the application of the
method also indicates problems with the expectations of theoreticians. First,
in contrast with the large theoretrical differences between the methods,
{\it Natural\/} is only marginally more useful than {\it Transient\/}, and neither of the two
blame assignment methods are highly superior to {\it Erasure\/}.  Second,
the cost of Transient's blame can be huge.  In turn, these
problems suggest that, on one hand, the existing theory does not properly
predict practice (i.e. the behavior of blame in real systems with real programs),
and on the other hand, the existing practice may need
additional experiments to collect more data.

>>>>

The paper's {\it second contribution\/} is a set of {\em results from
applying the evaluation method to three distinct blame assignment
strategies in approximately 72,000 different scenarios\/} (see
sec.~\ref{sec:results}): (1)~{\it Transient\/}, i.e. Reticulated's
tracking of typed/untyped boundaries~\citep{vss-popl-2017}; (2)~{\it
Natural\/}, i.e. Typed Racket's use of a higher-order contract system and
its blame assignment~\citep{ff-icfp-2002}; and (3) {\it Erasure\/}, i.e.
the approach of industrial systems, which forgo blame in favor of error
messages from the safety checks of the underlying language. The results
(see sec.~\ref{sec:discussion}) are at least somewhat surprising.  In
principle they validate the conjectures behind the work of theoreticians.
First, a blame assignment strategy helps with the search for
impedance mismatches between the specified types and the behavior of
untyped components.  Second, Natural's wrapper-based blame tracking is
more useful than Transient's ``collective blame'' tracking algorithm,
which in turn is superior to Erasure. {\em But\/}, the application of the
method also indicates problems with the expectations of theoreticians. First,
in contrast with the theoretical differences between the methods,
{\it Natural\/} is only marginally more useful than {\it Transient\/}, and neither of the two
blame assignment methods are highly superior to {\it Erasure\/}.  Second,
the cost of Transient's blame can be huge.  In turn, these
problems suggest that, on one hand, the existing theory does not properly
predict practice (i.e. the behavior of blame in real systems with real programs),
and on the other hand, the existing practice may need
additional experiments to collect more data.

### 2021-05-07 writing/mutate-benchmarks.tex
Touch up previous edits

\item vary in size, complexity and purpose;\footnote{A finer point worth noting is that since the rational programmer assumes observable changes in program behavior are due to its actions, the programs should be deterministic.}

@@

What exists is \citet{gtnffvf-jfp-2019}'s collection of Typed Racket programs
for systematically measuring the implementation's performance. The benchmark
suite consists of fully typed, correct programs, written by a number of
different authors, who have maintained and evolved these programs over several
years. The programs range widely in size, complexity, purpose, origin and in
programming style. The latter is mostly a reference to their reliance on many
Typed Racket features: occurrence typing~\citep{tf-icfp-2010}, types for mutable
and immutable data structures~\citep{hpst-sfp-2010}, types for first-class
classes and objects~\citep{tsdtf-oopsla-2012}, and types for Racket's numeric
tower~\citep{stathff-padl-12}.

>>>>

\item vary in size, complexity and purpose;

@@

What exists is \citet{gtnffvf-jfp-2019}'s collection of Typed Racket programs
for systematically measuring the implementation's performance. The benchmark
suite consists of fully typed, correct programs, written by a number of
different authors, who have maintained and evolved these programs over several
years. The programs range widely in size, complexity, purpose, origin and in
programming style. The latter is mostly a reference to their reliance on many
Typed Racket features: occurrence typing~\citep{tf-icfp-2010}, types for mutable
and immutable data structures~\citep{hpst-sfp-2010}, types for first-class
classes and objects~\citep{tsdtf-oopsla-2012}, and types for Racket's numeric
tower~\citep{stathff-padl-12}.
Finally, all of the programs are deterministic, so any changes in the programs' behavior
between runs can be solely attributed to the actions of the rational programmer.
